---
title: "Homework 3 solutions"
author: "Rio Yan"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

Let's make a table!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


Apples vs ice cream...

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```


## Problem 2
Load and tidy data. 
final dataset should include all originally observed variables and values; have useful variable names; 
include a weekday vs weekend variable; 
and encode data with reasonable variable classes. 
Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = 
  read_csv(
    "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity_count"
    ) %>% 
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>% 
  mutate(
    weekday_vs_weekend = 
      case_when(
        day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
        day %in% c("Saturday", "Sunday") ~ "weekend")
    ) %>% 
  mutate(
    week = factor(week),
    weekday_vs_weekend = factor(weekday_vs_weekend),
    day_id = factor(day_id),
    day = factor(day),
    day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday", "Sunday")),
    minute_of_day = as.numeric(minute_of_day)
    ) %>% 
  relocate(day_id, week, day, weekday_vs_weekend)

str(accel_df)
```
descriptio:::::


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  )

# We can see from the table that usually the participant has an average activity count around 400000. The participant's Tuesday to Thursday's activity count has a smaller variance compares to Monday, Saturday and Sunday. The first day of the first week has the least amount of activity count for that week. Week four and five's weekend have the least amount of activity. Saturdays in general have the least sum of activity count. 
```

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_df %>% 
  ggplot(aes(x = minute_of_day, y = activity_count, color = day)) +
  geom_point(alpha = .5) +
  geom_smooth()
```


## Problem 3
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

Load data
```{r}
data("ny_noaa")
```

The dataset contains data from New York state weather stations. This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns.

There are seven important variables in the dataset, weather station ID, date of observation, precipitation (tenths of mm), snowfall (mm), snow depth (mm), maximum and minimum temperature (tenths of degrees C). Each weather station may collect only a subset of these variables, and therefore the resulting dataset contains extensive missing data.


Do some data cleaning.  Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa_df = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day")) %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
  ) %>% 
  mutate(
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  )

ny_noaa_df %>% 
  count(snow) %>% 
  arrange(desc(n))
```
0 is the most commonly observed values. Because NY doesnt really snow...?


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

how to organize data so that max temp in jan and july
```{r}
ny_noaa_df %>% 
  group_by(id, year, month) %>%
  filter(month %in% c("01", "07")) %>% 
  summarize(
    mean_tmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id)) +
  geom_point() +
  geom_path() +
  facet_grid(. ~ month)

str(ny_noaa_df)
```

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 
```{r}

  
```



(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.



