---
title: "Homework 3 solutions"
author: "Rio Yan"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

Let's make a table!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


Apples vs ice cream...

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```


## Problem 2

Load and tidy data. 

```{r}
accel_df = 
  read_csv(
    "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity_count"
    ) %>% 
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>% 
  mutate(
    weekday_vs_weekend = 
      case_when(
        day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
        day %in% c("Saturday", "Sunday") ~ "weekend")
    ) %>% 
  mutate(
    week = factor(week),
    weekday_vs_weekend = factor(weekday_vs_weekend),
    day_id = factor(day_id),
    day = factor(day),
    day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday", "Sunday")),
    minute_of_day = as.numeric(minute_of_day)
    ) %>% 
  relocate(day_id, week, day, weekday_vs_weekend)
```

Description of new dataset:

The dataset contains five weeks of accelerometer data collected on a 63 year-old male with BMI 25. This dataset contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns, so there are `r nrow(accel_df)` observations in total.  "Day_id", "week", and "day" variables are day variables used to document and identify different study days. The new variable "weekday_vs_weekend" separates the study days into two categories, weekdays and weekends. The "minute_of_day" and "activity_count" store data from the original dataset activity_1 to activity_1440. "Minute_of_day" collects minute data for each day starting from the 1st minute to the last 1440th minute, and "activity_count" calculates how active the participant was during each minute. 



Aggregate across minutes to create a total activity variable for each day, and create a table showing these totals.
```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(digits = 1)
```

Description on trends:

We can see from the table that most of the activity count is around 400000, but there is no prominent trends in general. 
The participant's Tuesday to Thursday's activity count has a smaller variance compares to Monday, Saturday and Sunday. Saturdays have the largest variation in activity count, ranging from 1440 to 607175. Saturdays on week four and five have the least amount of activity count. It's also interesting to note that if the participants have an aggregated activity count over 600000, he will definitely have a significantly decrease in activity count the next day.



Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r}
accel_df %>% 
  ggplot(aes(x = minute_of_day, y = activity_count, color = day)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.4) +
  labs(title = "Activity across time plot",
       x = "Minute of day",
       y = "Activity count")
```
Describe in words any patterns or conclusions you can make based on this graph:



## Problem 3
To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

Load data
```{r}
data("ny_noaa")
```

The dataset contains data from New York state weather stations. This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns.

There are seven important variables in the dataset, weather station ID, date of observation, precipitation (tenths of mm), snowfall (mm), snow depth (mm), maximum and minimum temperature (tenths of degrees C). Each weather station may collect only a subset of these variables, and therefore the resulting dataset contains extensive missing data.


Do some data cleaning.  Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa_df = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day")) %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
  ) %>% 
  mutate(
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  )

ny_noaa_df %>% 
  count(snow) %>% 
  arrange(desc(n))
```
0 is the most commonly observed values. Because NY doesnt really snow...?


********Make a two-panel plot showing the average max temperature in January and in July in each station across years.

```{r}
ny_noaa_df %>% 
  group_by(id, year, month) %>%
  filter(month %in% c("01", "07")) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id, color = month)) +
  geom_line() +
  facet_grid(. ~ month) +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1)) +
  labs(title = "Mean average temperature for Jan and Jul across stations and years",
       x = "Year",
       y = "Average max temperature")
```
Is there any observable / interpretable structure? Any outliers?
***********


Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 
```{r}
#(i) plot
tmax_tmin_p = 
  ny_noaa_df %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex() +
  labs(title = "Max temperature vs min temperature(degree C)",
       x = "Maximum temperature",
       y = "Minimun temperature")
```

(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
#(ii) plot
snowfall_p = 
  ny_noaa_df %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes( x = as.factor(year), y = snow)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1)) +
  labs(title = "Snowfall values greater than 0 and less than 100 by year ",
       x = "Year",
       y = "Snowfall values")
```

two-panel plot combining the two
```{r}
library(patchwork)
tmax_tmin_p + snowfall_p
```


